---
title: "PostEddyPro Package Usage for CH4 Flux Gapfilling"
output: html_document
---

# Installing packages
```{r}
install.packages("devtools")
devtools::install_github("bravemaster3/PostEddyPro")
```

# Loading the required libraries and your dataset
```{r echo = FALSE}
library(PostEddyPro)
library(data.table)
library(dplyr)
library(caret)
df <- fread("data/EC_data_qc_filtered.csv", header = TRUE)
#VIEW DF TO ENSURE IT LOOKS GOOD, ESPECIALLY THE timestamp column
```

# Running the gapfilling
!!!Important to note that all the input parameters should not have gaps
This means that they have been gapfilled prior to this step.
```{r echo=FALSE}
# IF YOU WANT TO TEST A SMALL SUBSECTION, UNCOMMENT FIRST THIS AND RUN. THE ROW NUMBERS ARE ONLY FOR THE SAMPLE DATASET. then ofcourse change "rf_gapfiller_fast(site_df=df," below to  "rf_gapfiller_fast(site_df=df_subset,"
# df_subset <- df[1825:4014,]

# Define a predictor list
predictor_list <- c("Ta_f","Ts_f","WTD_f","Pa_f","PARin_f","PARout_f","RH_f","VPD_f","P_f","yearly_sin","yearly_cos","delta") 
gapfilling_list <- rf_gapfiller_fast(site_df=df,
                                     datetime="datetime",
                                     flux_col="ch4_flux_final", # column containing the filtered, i.e. quality checked methane flux column
                                       #List of predictors for the Random forest model. Change to what you have available. Note that "yearly_sin", "yearly_cos" and "delta" are calculated internally by the rf_gapfiller_fast function
                                     preds=predictor_list,
                                     max_mtry = NULL, #This is for trying values of mtry up to the max of either
                                     #sqrt of the number of predictors, or the 1/3rd of the number of predictors
                                     sitename="Halmyran" #This variable is simply for annotating the plots with the site name.
)
```
```{r echo=FALSE}
# Save the list if needed
saveRDS(gapfilling_list, "output/gf_rds.RDS")
# You can read it like this (uncomment to run it):
# gapfilling_list <- readRDS("output/gf_rds.RDS")
```

```{r echo=FALSE}
# You can visualize the predicted vs. measured from the cross validation
gapfilling_list$pred_meas
# Or the original vs. gapfilled after filling gaps on the model trained on the full available dataset
gapfilling_list$gf_meas_time
#Or view the final dataframe which includes the input given to the gapfilling function and the column "<flux_col>_filled", as well as the column "quality", that tells you which of the "<flux_col>_filled" values are original and which ones are gapfilled.
View(gapfilling_list$site_df)
# Or view the cross validation r² and rmse 
gapfilling_list$r_sq_cv
gapfilling_list$rmse_cv
# Or see the content of the list with 
str(gapfilling_list, max.level = 1)
# Where final model is the random forest model trained on all available dataset, which is what is used for filling the gaps after doing cross validation. Then, the tuningmodel contains information about the best hyperparameters obtained during the cross validation and tuning.
```


# That was the end of the gapfilling, but if you also want to try some montecarlo simulation for uncertainty estimation, then modify the following script to achieve that
# NOTE THAT HERE I named my timestamp column "datetime", and also, "ch4_flux_final" is my quality filtered column that I started with up in this code. if you have a different name, you will have to change that... "ch4_flux_final_filled" follows the same reasoning... some folders may need to be created, like the content of output (e.g., create "Montecarlo" inside output, etc.). Hopefully not too much to debug below.

```{r echo=False}
mc_sim_list <- montecarlo_sim(df_gf = gapfilling_list$site_df, #the gapfilled flux dataframe
                                   df_cv = ggplot2::ggplot_build(gapfilling_list$pred_meas)$plot$data, # This is not saved, but in the previous list, but the graph was from the cross validation data, and so ggplot_build rebuilds the dataframe of the cross validation from this graph
                                   flux_col="ch4_flux_final",#in df_cv # THIS COLUMN SHOULD BE YOUR quality filtered column, whatever you named it. Change the name here if you use a different name.
                                   flux_pred_col="ch4_flux_final_filled", #From cross validation (df_cv) if you quality filtered data was named "FCH4", then this will be named "FCH4_filled"
                                   datetime = "datetime", # Change to whatever you called your timestamp column
                                   preds = predictor_list, # Same predictors here as you used for gapfilling
                                   n = 100, # Number of montecarlo simulations you want to make
                                   saving_folder = "output/Montecarlo", #change path if you don't want it to just go to your output folder in this working directory. make sure this directory exists
                                   flux_sign = "positive" # this is because I use this function as well for CO2 fluxes, which can also be negative/positive
)


#Gapfilling montecarlo
montecarlo_sim_gf(mc_sim_path = "output/Montecarlo",
                  mtry = gapfilling_list$tuningmodel$bestTune$mtry, #this is the best tuning during the caret crossvalidation, should be fetched from the saved RF gapfilling model prior to using this function
                  preds = predictor_list, #same as used in the original gapfilling
                  flux_col = "ch4_flux_final", #the quality filtered column name (i.e. before gapfilling)
                  mc_sim_gf_path = "output/Montecarlo_gf") #  make sure this directory exists

#Merging gapfilled montecarlo
df_mc <- merge_montecarlo_sims(dir = "output/Montecarlo_gf",
                               saving_dir = "output/Montecarlo_gf_merged") # Make sure this directory exists

# df_mc <- fread("output/Montecarlo_gf_merged/monte_carlo_all.csv")

df_mc_sd <- df_mc %>% 
  group_by(datetime) %>% 
  summarise(gf_uncertainty = sd(ch4_flux_final_filled, na.rm = TRUE))

# This gives you the gapfilling uncertainty, which you can still aggregate with the random mesurement error (as sqrt(gapfilling_error^2 + random error^2)

#Preparing random measurement error
err_df <- gapfilling_list$site_df %>% 
  select(datetime, rand_err_ch4_flux, ch4_flux_final_filled, quality, all_of(predictor_list)) #This column "rand_err_ch4_flux" is from the eddypro output file, and should be in side_df if you have carried it over...

err_df$rand_err_ch4_flux[which(err_df$quality == "gapfilled")] <- NA
err_df$quality <- NULL
preProcValues <- preProcess(err_df, method = c("bagImpute"))
err_df_imp <- predict(preProcValues, err_df)

plot(rand_err_ch4_flux~datetime, data=err_df_imp)

# Then you can do the aggregation of rand_err_ch4_flux and 

# Aggregate gapfilling uncertainty with random measurement error
final_uncertainty <- err_df_imp %>%
  left_join(df_mc_sd, by = "datetime") %>%
  mutate(total_uncertainty = sqrt(rand_err_ch4_flux^2 + gf_uncertainty^2))
```